Observed vs. Simulated Year Comparison

### 2023-03-10 03-post-process/compare-sim-to-obs.Rmd
Use geometric mean for sim-v-obs ratio summaries
See #168 for complete details

@@

```{r ratio_plot fn}
ratio_plot = function(ratios, line_y, ...) {

@@

@@

```{r ratio_plots fn}
ratio_plots = function(x, main) {
  par(mfrow = c(1,3), oma = c(0,2,2,0))
  ratio_plot(x[,"sim-v-obs","cv",], main = "CV", line_y = 1)
  ratio_plot(x[,"sim-v-obs","mean",], main = "Mean", line_y = 1)
  ratio_plot(x[,"sim-v-obs","acf",], main = "ACF (Sim - Obs)", line_y = 0)
  mtext(side = 2, outer = TRUE, adj = 0.75, line = 1, "Ratio (Sim:Obs)")
  mtext(side = 3, outer = TRUE, line = 0.5, main, cex = 1.25, font = 2)
}
```

@@

```{r summarize_ratios fn}
summarize_ratios = function(x, Class, Type) {
  xsumm = as.data.frame(t(apply(x[,"sim-v-obs",,], 3, colMeans)))
  xsumm = cbind(Class = Class, Type = Type, ID = rownames(xsumm), xsumm); rownames(xsumm) = NULL
  return(xsumm)
}
```

@@

These plots show summaries of a quantity in the model:

1. **Inter-annual CV**: measures how variable the quantity is among years
2. **Inter-annual mean**: measures the central tendency of the quantity across years
3. **Lag-1 autocorrelation**: measures whether consecutive years are independent

@@

These summary statistics were calculated for each posterior sample. In all plots, the cross mark represents the posterior median, the thick bar represents the central 50% of posterior samples, and the thin bar represents the central 95% of posterior samples.

@@

These tables summarize the mean ratio values shown in the plots above, and are intended to be a quick visual for identifying areas where the simulated outcomes are systematically different in some way (central tendency, interannual variability, or temporal autocorrelation) from the observed outcomes. The number in each cell represents the posterior mean value of:

* **Across Year Mean** is the ratio `mean(simulated years)/mean(observed years)`. Values greater than 1 indicate the value in simulated years is larger in magnitude than in observed years.
* **Across Year CV** is the ratio `cv(simulated years)/cv(observed years)`. Values greater than 1 indicate the variability from one year to the next is larger larger than in observed years.
* **Lag-1 Autocorrelation** is the difference `acf(simulated years) - acf(observed years)`. Values greater than 0 indicate the quantity shows more autocorrelation in the simulated years than in observed years.

>>>>

```{r geomean fn}
geomean = function(x) exp(mean(log(x)))
```

@@

```{r ratio_plot fn}
ratio_plot = function(ratios, line_y, pt_fn = mean, ...) {

@@

  # replace median with output of supplied point estimate function
  summs["50%",] = apply(ratios, 2, pt_fn)

@@

```{r ratio_plots fn}
ratio_plots = function(x, main) {
  par(mfrow = c(1,3), oma = c(0,2,2,0))
  ratio_plot(x[,"sim-v-obs","cv",], main = "CV", pt_fn = geomean, line_y = 1)
  ratio_plot(x[,"sim-v-obs","mean",], main = "Mean", pt_fn = geomean, line_y = 1)
  ratio_plot(x[,"sim-v-obs","acf",], main = "ACF (Sim - Obs)", pt_fn = mean, line_y = 0)
  mtext(side = 2, outer = TRUE, adj = 0.75, line = 1, "Ratio (Sim:Obs)")
  mtext(side = 3, outer = TRUE, line = 0.5, main, cex = 1.25, font = 2)
}
```

@@

```{r summarize_ratios fn}
# function to summarize a matrix of ratios
col_summs = function(x) {
  c(cv = geomean(x[,"cv"]), mean = geomean(x[,"mean"]), acf = mean(x[,"acf"]))
}

summarize_ratios = function(x, Class, Type) {
  xsumm = as.data.frame(t(apply(x[,"sim-v-obs",,], 3, col_summs)))
  xsumm = cbind(Class = Class, Type = Type, ID = rownames(xsumm), xsumm); rownames(xsumm) = NULL
  return(xsumm)
}
```

@@

These plots show the distribution of a quantity in the model:

1. **Inter-annual CV**: measures how variable the quantity is among years (uses arithmetic mean)
2. **Inter-annual mean**: measures the central tendency of the quantity across years (uses arithmetic mean)
3. **Lag-1 autocorrelation**: measures whether consecutive years are independent

@@

These summary statistics were calculated for each posterior sample. In estimate plots, the cross mark represents the posterior median, the thick bar represents the central 50% of posterior samples, and the thin bar represents the central 95% of posterior samples. The ratio plots have the same interpretation, except the cross mark represents the posterior geometric mean for the "CV" and "Mean" panels, and the posterior arithmetic mean for the "ACF" panels. The geometric mean is used to summarize central tendency of ratios across posterior samples because it is more appropriate for calculations based on multiplicative, rather than additive, variability (see issue [#168](https://github.com/bstaton1/GR-sslcm/issues/168) for more details).

@@

These tables summarize the ratio values shown in the plots above, and are intended to be a quick visual for identifying areas where the simulated outcomes are systematically different in some way (central tendency, interannual variability, or temporal autocorrelation) from the observed outcomes. The number in each cell represents the posterior mean (geometric for across year mean and across year CV; arithmetic for lag-1 autocorrelation) value of:

* **Across Year Mean** is the ratio `mean(simulated years)/mean(observed years)`. Values greater than 1 indicate the value in simulated years is larger in magnitude than in observed years.
* **Across Year CV** is the ratio `cv(simulated years)/cv(observed years)`. Values greater than 1 indicate the variability from one year to the next is larger than in observed years.
* **Lag-1 Autocorrelation** is the difference `acf(simulated years) - acf(observed years)`. Values greater than 0 indicate the quantity shows more autocorrelation in the simulated years than in observed years.
